---
title: "Parameter estimation with the ApsimX crop Model: a simple case"
output: html_document
params:
  apsimx_path: !r  system("which Models", intern = TRUE)
  path_results: "/home/plecharpent/tmp/tests_SticsOptimizR/ApsimX_param_est_simple"
---


```{r setup, echo=FALSE}

# set global chunk options: 
library(knitr)
opts_chunk$set(cache=TRUE, autodep = TRUE)
dep_auto()
```

## Study Case

A simple parameter estimation with a single situation, a single observed variable and 2 estimated parameters, just to illustrate how to use the package.

> TODO : Data comes from a ... experiment (see description in apsimx file / interface...).

The parameter estimation is performed using the Nelder-Meade simplex method implemented in the `nloptr` package.

To regulate the number of simulations (and thus the duration of execution), reduce the number of repetition of the minimization and/or the maximum number of evaluation (see section "Set options for the parameter estimation method").

The execution of this Rmarkdown document requires the path to the ApsimX model. By default it uses the one calculated in this document header stored in `params$apsimx_file`
```{r}

apsimx_path <- params$apsimx_path
apsimx_path

```

If you want to change the value of this parameter you may use Knit, then *"Knit with parameters"* or set it manually in a chunk (`apsimx_path <- path/to/exe/file`) .

## Initialisation step

```{r setup and initializations, message=FALSE, results=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install and load the needed libraries
if(!require("ApsimOnR")){
  devtools::install_github("ApsimOnR")
  library("ApsiOnR")
}
if(!require("SticsOptimizR")){
  devtools::install_github("SticsRPacks/SticsOptimizR")
  library("SticsOptimizR")
}
if(!require("dplyr")){
  install.packages("dplyr",repos="http://cran.irsn.fr")
  library("dplyr")
}
if(!require("nloptr")){
  install.packages("nloptr",repos="http://cran.irsn.fr")
  library("nloptr")
}
if(!require("DiceDesign")){
  install.packages("DiceDesign",repos="http://cran.irsn.fr")
  library("DiceDesign")
}
if(!require("doParallel")){
  install.packages("doParallel",repos="http://cran.irsn.fr")
  library("doParallel")
}




```

## List of situations and variables to consider

```{r  eval=TRUE, message=FALSE, warning=FALSE}

sit_name="GattonRowSpacingRowSpace25cm"  # among "GattonRowSpacingRowSpace25cm", "GattonRowSpacingRowSpace50cm","GattonRowSpacingRowSpaceN0"

var_name = c("Wheat.Leaf.LAI","Clock.Today") # or "Wheat.AboveGround.Wt"

```


## Run the model before optimization for a prior evaluation

In this case, the argument `param_values` of the wrapper is not set: the values of the model input parameters are all read in the model input files.

```{r eval=TRUE, message=FALSE, warning=FALSE}


# Set the model options (see '? apsimx_wrapper_options' for details)
#
# Normally apsimx_path has been set in Rmarkdown doc header (see params list).
# But it can be overrided here, or using the feature "Knit with parameters"
# For a linux OS : apsimx_path="/usr/local/bin/Models"
# For a Windows OS: apsimx_path="C:/path/to/apsimx/bin/Models.exe"
apsimx_path <- params$apsimx_path

files_path <- system.file(file.path("extdata","apsimx_files"),package = "ApsimOnR")
apsimx_file <- file.path(files_path, "template.apsimx")

# Setting met files path
met_files_path <- files_path

# Setting observed data files path
obs_files_path <- files_path

# Setting sqlite db tables names 
predicted_table_name <- "DailyReport"
observed_table_name <- "Observed"

model_options=apsimx_wrapper_options(apsimx_path = apsimx_path,
                                     apsimx_file =  apsimx_file,
                                     variable_names = var_name,
                                     predicted_table_name = predicted_table_name,
                                     met_files_path = met_files_path,
                                     observed_table_name = observed_table_name,
                                     obs_files_path = obs_files_path) 

# Run the model (on all situations found in the apsimx_file)
sim_before_optim=apsimx_wrapper(model_options=model_options)


```

## Read and select the corresponding observations

We only keep observations for situation `sit_name` and variable `var_name` (`obs_list` define the list of situations and variables that will be used in the parameter estimation process).

```{r eval=TRUE, message=FALSE, warning=FALSE}
# At the moment, observed data are read from the db file after the first simulation ran before optimization.
#But they may be loaded using the original xlsx data file (from the files_path)  

obs_list <- read_apsimx_output(sim_before_optim$db_file_name,
                               model_options$observed_table_name,
                               model_options$variable_names,
                               names(sim_before_optim$sim_list))

sit_name="GattonRowSpacingRowSpace25cm"

obs_list=obs_list[sit_name]
names(obs_list) <- sit_name

obs_list

```

## Set prior information on the parameters to estimate

`prior_information` determines the list of parameters that will be estimated in the parameter estimation process and associated bounds information (only uniform distributions for the moment, others will come soon).

```{r eval=TRUE, message=FALSE, warning=FALSE}
# 2 parameters here: ExtinctionCoeff and RUE, of prior distributions U([0.4,0.6]) and U([1.4,1.6])
# This is also used to set the list of parameters to estimate
prior_information <- 
  list(lb=c(.Simulations.Replacements.Wheat.Leaf.ExtinctionCoeff.VegetativePhase.FixedValue=0.4,
            .Simulations.Replacements.Wheat.Leaf.Photosynthesis.RUE.FixedValue=1.4),
       ub=c(.Simulations.Replacements.Wheat.Leaf.ExtinctionCoeff.VegetativePhase.FixedValue=0.6,
            .Simulations.Replacements.Wheat.Leaf.Photosynthesis.RUE.FixedValue=1.6))

```

## Set options for the parameter estimation method

`optim_options` should contain the options of the parameter estimation method.
Here we defined a few options for the simplex method of the `nloptr` package (defautl method in main_optim).
The full set of options for the simplex method can be found in the [vignette of nloptr package](https://cran.r-project.org/web/packages/nloptr/vignettes/nloptr.pdf).

The number of repetitions `nb_rep` is advised to be set at least to 5, while 10 is a reasonable maximum value.
`maxeval` should be used to stop the minimization only if results have to be produced within a given duration, otherwise set it to a high value so that the minimization stops when the criterion based on the relative tolerance `xtol_rel` is satisfied. 

```{r eval=TRUE, message=FALSE, warning=FALSE}

optim_options=list() 
optim_options$nb_rep <- 2 # Number of repetitions of the minimization 
# (each time starting with different initial
# values for the estimated parameters) 
optim_options$maxeval <- 10 # Maximum number of evaluations of the 
# minimized criteria 
optim_options$xtol_rel <- 1e-05 # Tolerance criterion between two iterations
# (threshold for the relative difference ofparameter values between the 2 previous 
# iterations)
#
# path of results directory
# Normally path_results has been set in Rmarkdown doc header (see params list).
# But it can be overrided here, or using the feature "Knit with parameters"
#  optim_options$path_results <- "/path/to/results/dir"
optim_options$path_results <- params$path_results

#optim_options$ranseed <- 1234 # random seed  


```

## Run the optimization

The **Nelder-Meade** simplex is the default method. So for the moment, no need to set the `optim_method` argument; it is the only method interfaced (others will come soon).
Same for `crit_function`: a value is set by default (`concentrated_wss`, see Wallach *et al.*, 2011). For the moment it is the only  proposed criterion (others will come soon). The user can implement its own criterion (see inputs and outputs required in the `concentrated_wss` function).

```{r eval=TRUE, message=FALSE, warning=FALSE}

param_est_values=main_optim(obs_list=obs_list,
                            crit_function=concentrated_wss,
                            model_function=apsimx_wrapper,
                            model_options=model_options,
                            optim_options=optim_options,
                            prior_information=prior_information)

print(paste("Results of the optimization were saved in",
            optim_options$path_results," folder."))



```

The results printed in output on the R console are the following:
```{r echo=TRUE, eval=TRUE}
# ExtinctionCoeff and RUE
param_est_values$final_values


```

Complementary graphs and data are stored in the **`r optim_options$path_results`** folder. Among them, the `EstimatedVSinit.pdf` file contains the following figures: 

```{r echo=FALSE, image_grobs, fig.show = "hold", out.width = "45%", fig.align = "default"}

#print("ALL data from main_optim...")

#print("init values")
init_values <- param_est_values$init_values
#init_values
#print("est values")
est_values <- param_est_values$est_values
#est_values
#print("final values")
final_values <- param_est_values$final_values
#final_values
param_names <- names(param_est_values$final_values)
#print("min index")
ind_min_crit <- param_est_values$ind_min_crit
#ind_min_crit

bounds=get_params_bounds(prior_information)

for (ipar in 1:length(param_names)) {
  plot(init_values[,ipar], est_values[,ipar],
       main = "Estimated vs Initial values of the parameters for different repetitions",
       text(init_values[,ipar], est_values[,ipar], pos=1,col="black"),
       xlim = c(bounds$lb[ipar],bounds$ub[ipar]),
       ylim = c(bounds$lb[ipar],bounds$ub[ipar]),
       xlab = paste("Initial value for", param_names[ipar]),
       ylab = paste("Estimated value for", param_names[ipar]))
  text(init_values[ind_min_crit,ipar], est_values[ind_min_crit,ipar],
       labels = ind_min_crit, pos=1,col="red")
}

```

Figure 1:  plots of estimated vs initial values of parameters *ExtinctionCoeff* and *RUE*. Numbers represent the repetition number of the minimization. The number in red, 2 in this case, is the minimization that lead to the minimal value of the criterion among all repetitions. In this case, minimizations converge towards 2 different values for one of the *RUE* parameter , which indicates the presence of a local minimum. Values of *ExtinctionCoeff* are relatively close to the lower bound value. In realistic calibration cases this may indicate the presence of a large error in the observation values or in the model output values simulated (this simple case with only one situation does not allow to derive such conclusion).

```{r echo=FALSE }
#load(file.path(optim_options$path_results,"optim_results.Rdata"))
nlo <- param_est_values$nlo
last_rep <- length(nlo)
```


The **`r optim_options$path_results`** folder also contains the `optim_results.Rdata` file that store the nlo variable, a list containing the results of the minimization for each repetition. Here after are displayed the results for the repetition `r last_rep`:

```{r echo=TRUE, eval=TRUE}
nlo[[last_rep]]

```

## Run the model after optimization

In this case, the `param_values` argument is set so that estimated values of the parameters overwrite the values defined in the model input file ('.apsimx`).

```{r eval=TRUE, message=FALSE, warning=FALSE}

sim_after_optim=apsimx_wrapper(param_values=param_est_values$final_values,
                               model_options=model_options)

```

## Plot the results

```{r eval=TRUE, message=FALSE, warning=FALSE,out.width = '80%'}
var_name = c("Wheat.Leaf.LAI")

par(mfrow = c(1,2))

# Simulated and observed LAI before optimization
Ymax=max(max(obs_list[[sit_name]][,var_name], na.rm=TRUE),
         max(sim_before_optim$sim_list[[sit_name]][,var_name], na.rm=TRUE))
plot(sim_before_optim$sim_list[[sit_name]][,c("Date",var_name)],type="l",
     main="Before optimization",ylim=c(0,Ymax+Ymax*0.1))
points(obs_list[[sit_name]]$Date,obs_list[[sit_name]][[var_name]],col="red")
plot(sim_after_optim$sim_list[[sit_name]][,c("Date",var_name)],type="l",
     main="After optimization",ylim=c(0,Ymax+Ymax*0.1))
points(obs_list[[sit_name]]$Date,obs_list[[sit_name]][[var_name]],col="red")


```

```{r echo = FALSE, out.width = '80%', fig.cap="Figure 2: plots of simulated and observed target variable before and after optimization. If the gap between simulated and observed values has been drastically reduced (even) for some of the points: the minimizer has done its job!"}
# exporting plot to a png file
png <- dev.print(device = png,
          file = file.path(optim_options$path_results,"sim_obs_plots.png"),
          width = 15,
          height = 10,
          units = "cm",
          res=1000, )

```




